---
title: "IS607 Project 1"
author: "Aadi Kalloo"
date: "Due February 28, 2016"
output: html_document
---

Load data from Github, place into _chessinfo_, and initialize _chessdata_ dataframe:

```{r}
library(stringr)
chessinfo = suppressWarnings(readLines("https://raw.githubusercontent.com/aadikalloo/AadiMSDA/master/IS607%20Data%20Acquisition%20and%20Management/Project-1/tournamentinfo.txt"))
chessdata = as.data.frame(1)
```


Extract names from _chessinfo_ and store in _chessdata_ dataframe. Also remove extra spacing from names:

```{r}
row = 0
count = 0

for (count in seq(5, 196, 3)) {
  row = row + 1
  temp_loc = regexpr("[A-Z].+?\\|", chessinfo[count])
  chessdata[row, 1] = as.character(substr(chessinfo[count], temp_loc, attr(temp_loc, "match.length")))
  
}

count = 0
for (count in 1:length(chessdata[, 1])) {
  temp_loc = gregexpr("[A-Z] ", chessdata[count, 1])
  chessdata[count, 1] = substr(chessdata[count, 1], 1, temp_loc[[1]][length(temp_loc[[1]])])
}
```



Extract player state, total points, and player's pre-tournament rating:

```{r}
row = 0
count = 0
for (count in seq(6, 196, 3)) {
  row = row + 1
  temp_loc = regexpr(" [A-Z].+?\\|", chessinfo[count])
  chessdata[row, 2] = as.character(substr(chessinfo[count], temp_loc, attr(temp_loc, "match.length")))
}

row = 0
count = 0
for (count in seq(5, 196, 3)) {
  row = row + 1
  chessdata[row, 3] = as.numeric(str_extract(chessinfo[count], "[[:digit:]]\\.[[:digit:]]+"))
}

row = 0
count = 0
for (count in seq(6, 196, 3)) {
  row = row + 1
  temp_loc = 
  chessdata[row, 4] = as.integer(str_extract_all(chessinfo[count], "(\\d)+")[[1]][2])
}


```

Get list of opponents, find each opponent's pre-tournament rating, and create average score of opponents' pre-tournament ratings. Not all players face the same number of opponents and this is accounted for by removing NA's:

```{r}
row = 0
count = 0
for (count in seq(5, 196, 3)) {
  row = row + 1
  opponents = as.numeric(str_extract_all(chessinfo[count], "(\\d)+")[[1]])[4:10]
  opponents = opponents[!is.na(opponents)]
  
  sum = 0
  for (opnum in opponents) {
    sum = sum + chessdata[opnum, 4]
  }  
  chessdata[row, 5] = as.integer(sum/length(opponents))
}
```

Create column names:

```{r}
names(chessdata) = c("Player_name", "Player_state", "Total_points", "Pre_rating", "Avg_Opponent_pre_rating")
```

Now to take a look at the data:

```{r}
str(chessdata)
head(chessdata)
summary(chessdata)

```

