---
title: "IS605 Assignment 1"
author: "Aadi Kalloo"
date: "Due February 7, 2016"
output: html_document
---
### Problem Set 1

**Question 1:**
```{r}
u <- c(0.5, 0.5) #or rep(0.5, 2)
v <- c(3, -4)
uvdot = u %*% v
print(paste0("The dot product is: ", uvdot))
```

**Question 2:**
```{r}
vec_len <- function(vec_in) {sqrt(sum((vec_in^2))) }
u_len = vec_len(u)
v_len = vec_len(v)
print(paste0("The magnitude of u: ", sprintf("%.3f", u_len)))
print(paste0("The magnitude of v: ", sprintf("%.3f", v_len)))
```

**Question 3:**
```{r}
3*u - 2*v
```

**Question 4:**
```{r}
cosA = uvdot/(u_len*v_len)  #cos of Angle
A = acos(cosA)
print(paste0("The angle between u and v is ", sprintf("%.3f", A), " radians"))
```

       
### Problem Set 2
      
First define the matricies:   
```{r}
coef_mat = matrix(c(1,2,-1,1,-1,-2,3,5,4), nrow=3)
constraint_mat = matrix(c(1,2,6), nrow=3)

A = coef_mat
b_vec = constraint_mat
```

Now two functions are created: *createTriangular* returns the upper triangular matrix;  *backSubstitution* uses the upper triangular and the transformed b vector to return the solutions to the 3x3 matrix. *solveMatrix* puts everything together into one neat function:

```{r}
createTriangular <- function(A_matrix, bvec) {  
  nrows = length(A_matrix[, 1])
  ncols = length(A_matrix[1, ])
  
  for (a in 2:nrows) {
    multiplier = 0

    for (b in 1:(a-1)) {
        rows_back = 1

        pivot = A_matrix[a, b]
        val2 = A_matrix[a-rows_back, b]
        if (val2 == 0) {rows_back = 2; val2 = A_matrix[a-rows_back, b] }
        multiplier = pivot/val2
        A_matrix[a,] = A_matrix[a,] - A_matrix[a-rows_back,]*multiplier
        bvec[a,] = bvec[a,] - bvec[a-rows_back,]*multiplier
    }
  }

  return(list(A_matrix, bvec))
}

backSubstitution <- function(mat1, b) {
  solutions = c(0,0,0)
  for (num in length(b):1) {
    solutions[num] = (b[num] - mat1[num, 3]*solutions[3] - mat1[num, 2]*solutions[2])/mat1[num, num]
  }
  solutions = matrix(solutions, nrow=length(b))
  return(solutions)
}

solveMatrix <- function(A_matrix, bvec) {
  
  U_matrix_list = createTriangular(A_matrix, b_vec)
  U_matrix = U_matrix_list[1][[1]]
  #print(U_matrix)     #Print the Upper Triangular form matrix
  b_transformed = U_matrix_list[2][[1]]
  #print(b_transformed)     #Print the transformed b vector matrix
  solutions = backSubstitution(U_matrix, b_transformed)
  
  return(solutions)
}
```

Call the functions to see solutions to x vector: 
```{r}
xsolutions = solveMatrix(A, b_vec)
print(xsolutions)
```

Check with native solver: 
```{r}
solve(A, b_vec)
```

Test code versatility with another 3x3 matrix. This one starts with a *zero* pivot in the second row:
```{r}
coef_mat = matrix(c(3,0,4,1,1,2,4,2,1), nrow=3)
constraint_mat = matrix(c(3,1,4), nrow=3)

A = coef_mat
b_vec = constraint_mat
print(A)
print(b_vec)

xsolutions = solveMatrix(A, b_vec)
print(xsolutions)

solve(A, b_vec)
```


