---
title: "IS607 Week 8"
author: "Aadi Kalloo"
date: "Due March 20, 2016"
output: html_document
---

This assignment focuses on parsing information from html, xml, and json files. To do so, the RCurl, jsonlite, and htmltable (not to be confused with htmlTable) packages are used:


```{r}
options(warn = -1)
#install.packages("htmltab")
library(htmltab)
library(RCurl)
library(jsonlite)
library(XML)
library(knitr)
```

A small function to deal with multiple authors:

```{r}
reorganizeTable <- function(table1) {
for (i in 1:length(strsplit(table1$Authors, ","))) {
  authors = strsplit(table1$Authors, ",")
  if(length(authors[[i]]) > 1) {
    if (i < length(table1[ ,1])) { table1[i + 1:length(table1[ ,1]) + 1, ] = table1[i + 1:length(table1[ ,1]), ] }
    table1[i + 1, ] = table1[i, ]
    rownames(table1) = NULL
    for (j in 1:length(authors[[i]])) {
      table1$Authors[i+j-1] = authors[[i]][j]
      }
    }
}
  table1 = table1[-length(table1[ ,1]), ]
  return(table1)
}
```


A small dataset using cosmology texts was created for this assigment. The HTML table data is extracted here: 
```{r}
url = "https://raw.githubusercontent.com/aadikalloo/AadiMSDA/master/IS607%20Data%20Acquisition%20and%20Management/Books.html"
tableA = htmltab(doc = url)
tableA = data.frame(tableA, row.names = NULL)
tableA_2 = reorganizeTable(tableA)
```

Here we have the original table:
```{r}
kable(tableA)
```

Here the entry with multiple authors has been separated into different records:
```{r}
kable(tableA_2)
```


XML data is extracted here:
```{r}
url = "https://raw.githubusercontent.com/aadikalloo/AadiMSDA/master/IS607%20Data%20Acquisition%20and%20Management/Books.xml"
data = getURL(url, ssl.verifypeer=0L, followlocation=1L)
doc = xmlParse(data)
table3 = xmlToList(doc)
tableB = as.data.frame(matrix(unlist(table3), nrow = 3, ncol = 4, byrow=TRUE))
names(tableB) = names(tableA)
```

JSON data is extracted here:
```{r}
url = "https://raw.githubusercontent.com/aadikalloo/AadiMSDA/master/IS607%20Data%20Acquisition%20and%20Management/Books_json1.json"
my.JSON <- fromJSON(url)
tableC = as.data.frame(matrix(unlist(my.JSON), nrow = 3, ncol = 4, byrow=FALSE))
names(tableC) = names(tableA)
```

A comparison of the "untreated" HTML, XML, and JSON tables, respectively:
```{r}
kable(tableA)
kable(tableB)
kable(tableC)
```

It can be seen that all three dataframes are the same. 