---
title: "IS607 Assignment 1"
author: "Aadi Kalloo"
date: "January 31, 2016"
output: html_document
---

For this assignment, the Thyroid Disease Data Set from the UCI Repository is used.      
To clean the data, I import the data set and insert the variable names.     

```{r}
library(RCurl)
data_url <- getURL("https://archive.ics.uci.edu/ml/machine-learning-databases/thyroid-disease/sick.data")
data <- read.csv(text = data_url, header = FALSE)
names(data) <- c("age","sex","taking_thyroxine","query_thyroxine","antithyroid_med","sick","pregnant","thyroid_surgery","I131","query_hypothyroid","query_hyperthyroid","lithium","goitre","tumor","hypopituitary","psych","TSH_measured","TSH","T3_measured","T3","TT4_measured","TT4","T4U_measured","T4U","FTI_measured","FTI","TBG_measured","TBG","referral_source","status")

```
     
Variables with only 1 level should be removed: TBG and TBG_measured. Since all participants have the same value, these variables are unlikely to be useful.
```{r}
head(data)
str(data)
```

```{r,}
data$TBG <- NULL
data$TBG_measured <- NULL
```

The *status* column needs to be cleaned up. It is converted into two columns, one with disease status and the other containing the class:

```{r}
data$status <- strsplit(as.character(data$status), "\\.\\|")
data$classes <- lapply(data$status,"[",2)
data$status <- lapply(data$status,"[",1)
data$status <- as.factor(as.character(data$status))
data$classes <- as.numeric(as.character(data$classes))
```

I am interested in seeing data for males under 40 years of age who are negative for thyroid disease. Furthermore, I am only interested in variables that contain numerical data as these would be useful for calculations. This first involves finding the variables that contain numeric data and re-integrating back into the factor dataframe:

```{r}
data_chr <- data.frame(lapply(data, as.character), stringsAsFactors=FALSE)
num_data <- suppressWarnings(data.frame(sapply(data_chr[,grepl("^\\d", data_chr[1,])],as.numeric)))
data[names(num_data)] <- num_data
data_subsetted <- data[, sapply(data, is.numeric)]
data_subsetted$gender <- data$sex
remove(data_chr)
remove(num_data)
```

View the cleaned *data* dataframe: 
```{r}
str(data)
head(data)
```


View the subsetted data:

```{r}
str(data_subsetted)
head(data_subsetted)
summary(data_subsetted)
```



