---
title: "IS607 Assignment 2"
author: "Aadi Kalloo"
date: "February 8, 2016"
output: html_document
---



```{r}
library(RCurl)

pop_data_url <- getURL("https://raw.githubusercontent.com/aadikalloo/AadiMSDA/master/IS607%20Data%20Acquisition%20and%20Management/Assignment2_population.csv")
tb_data_url <- getURL("https://raw.githubusercontent.com/aadikalloo/AadiMSDA/master/IS607%20Data%20Acquisition%20and%20Management/tb_db.csv")

pop_data <- read.csv(text = pop_data_url, header = TRUE)
tb_data <- read.csv(text = tb_data_url, header = FALSE)
names(tb_data) <- c("country","year","sex","child","adult","elderly")

```

Let's take a look at the *pop_data* data frame: 

```{r}
head(pop_data)
summary(pop_data)
str(pop_data)
```

Let's take a look at the *tb_data* data frame: 

```{r}
head(tb_data)
summary(tb_data)
str(tb_data)
```

The goal is to create a data frame with the following 3 columns: Country, Year, Rate.

Re-code missing data:
```{r}
tb_data[tb_data == -1] = NA
```


Will need to combine age data into gender-stratified totals. Then combine gender data to get annual totals. 

First age data:
```{r}
  tb_data$agemerged = (tb_data$child + tb_data$adult + tb_data$elderly)
```

Now gender data:

```{r}
for (counter in seq(2, 3800, 2)) {
  tb_data$gender_merged[counter] =  sum(c(tb_data$agemerged[counter-1], tb_data$agemerged[counter]), 
                                        na.rm = TRUE)
}
```

Data now organized into a cleaner data frame of country, year, and total cases: 

```{r}
tb = data.frame( tb_data$country      [!is.na(tb_data$gender_merged)], 
                 tb_data$year         [!is.na(tb_data$gender_merged)], 
                 tb_data$gender_merged[!is.na(tb_data$gender_merged)] )

names(tb) <- c("Country", "Year", "Total Cases")
```

It is now very easy to create a data frame containing Country, Year, and Rate. 
Rate is expressed per hundred thousand:

```{r}
final_data = tb
names(final_data) = c("Country", "Year", "Rate")
final_data[final_data == 0] = NA
final_data <- final_data[order(final_data$Year),]
final_data$Rate = (final_data$Rate / pop_data$population) * 10^5
final_data$Rate = round(final_data$Rate, digits = 3)
```

Preview the *final_data* data frame:

```{r}
head(final_data)
str(final_data)
summary(final_data)
```

